Title

### 2013-04-07 TCAD2013/overclocking_tcad.tex
Modified CSA v3

If the input signal of a circuit is $k$ bits, truncation error occurs when the input signal is truncated from $k$ bits to $n$ bits. Under this premise, the mean value of the truncated bits at signal input ($E_{Tin}$) is given by (\ref{ET at single input}).
%
%\vspace{-1ex}
\begin{eqnarray}\label{ET at single input}
%\small
%\footnotesize
%\scriptsize
  E_{Tin}=\frac{1}{2}\sum_{i=n+1}^{k}2^{-i}=2^{-n-1}-2^{-k-1}
%  \vspace{-1ex}
\end{eqnarray}
%\normalsize

@@

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Generation of Overclocking Error}\label{subsub:Generation of Overclocking Error}
For a given $T_S$, the maximum length of error-free carry propagation is described by~(\ref{Max.length of carry chain}), where $f_S$ denotes the sampling frequency.
%
%\vspace{-1ex}
\begin{eqnarray}\label{Max.length of carry chain}
%\small
%\footnotesize
%\scriptsize
    b:=\left\lceil \frac{T_S}{\mu} \right\rceil=\left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}
%\normalsize

@@

$S_i$ and $\hat{S_i}$ can be determined using the equations from the previous statements in Section~\ref{subsub:Generation of Overclocking Error}. In the error-free case, the carry will propagate from bit $S_t$ to bit $S_{t+m-1}$, and we will obtain $S_{t+b}=S_{t+b+1}=\cdots=S_{t+m-2}=0$ for carry propagation, and $S_{t+m-1}=1$ for carry annihilation. However, when a timing violation occurs, the carry will not propagate through all these bits. Substituting these values into (\ref{Output Error}) yields (\ref{etm}). Interestingly, the value of overclocking error has no dependence on the length of carry chain $m$.
%
%Instead, $\hat{S}_{t+b}=\hat{S}_{t+b+1}=\cdots=\hat{S}_{t+m-2}=1$ and $\hat{S}_{t+m-1}=0$. It is worth noting that these expressions are obtained based on the assumption that all the internal bits are initialized to zero.
%
\begin{eqnarray}\label{etm}
%\small
%\footnotesize
  e_{tm}=\frac{\left|2^{t+m-1}-2^{t+m-2}-\dots-2^{t+b}\right|}{2^n}=2^{t+b-n}
\end{eqnarray}
%\normalsize

@@

\subsection{Area Overhead}
We then compare different adder structures in terms of the maximum word-length for a given operating frequency and the corresponding area consumption. Fig.~\ref{CSA3stage Timing} illustrates the maximum word-lengths for RCA and CSA with 2 stages and 3 stages under a range of operating frequencies. We only investigate 3 stages as the maximum stages number predicted in (\ref{CSA_StageNumber}) is 3. However, similar analysis can be performed on CSA with any stage numbers and the overall conclusion will still hold. It can be seen from Fig.~\ref{CSA3stage Timing} that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. This corresponds to smaller truncation errors. RCA only outperforms than CSA when very high frequency is applied. In addition, the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain, as expected.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
%  \caption{Maximum word-length for RCA and CSA across a variety of frequencies.}
%  \label{CSA3stage Timing}
%\end{figure}

@@

However, the accuracy benefits brought by CSA comes with the cost of large area overhead. Fig.~\ref{CSA3stage Area} depicts the number of Look-Up Tables (LUTs) in the FPGA used for all three structures. It can be seen that for all frequency values, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the number of the 2-stage CSA is $1.7\times\sim3.1\times$. This finding poses a question of whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
To answer this question, the silicon area is incorporated into the trade-offs between accuracy and performance. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision loss might results in large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. In the conventional design scenario, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking. For a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the minimum error expectation is the optimum design;
  \item If multiple designs achieve the same error expectation, then the design with minimum area is the optimum design;
  \item If the error expectation and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

The optimum design method for a variety of frequency and area constraints is demonstrated in Fig.~\ref{Tradeoff}, from which several observations can be made. If the available area is large enough, CSA serves as the optimum design option because it inherently operates faster than RCA, and it can be implemented with full precision with unlimited hardware resources. The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whilst RCA still keeps full precision. In this case, area instead of timing becomes the dominate factor. This leads to precision loss for CSA initially. Therefore we see RCA with overclocking is the optimum design method across almost the whole frequency domain. For an even stringent area constraint, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. Meanwhile, CSA with high stage numbers might not even be implemented at all.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology for a variety frequency and area constraints.}
  \label{Tradeoff}
\end{figure}

%We also notice that for small frequencies, RCA with either overclocking or truncation of word-length of the input signal is the best design choice for any area constraint, since

>>>>

If the input signal of a circuit is $k$ bits, truncation error occurs when the input signal is truncated from $k$ bits to $n$ bits. Under this premise, the mean value of the truncated bits at signal input ($E_{Tin}$) is given by (\ref{ET at single input}).
%
%\vspace{-1ex}
\begin{eqnarray}\label{ET at single input}
%\small
%\footnotesize
%\scriptsize
  \begin{split}
    E_{Tin} &= \frac{1}{2}\sum_{i=n+1}^{k}2^{-i}\\
    &= 2^{-n-1}-2^{-k-1}
%  \vspace{-1ex}
  \end{split}
\end{eqnarray}
%\normalsize

@@

\subsection{Probabilistic Model of Overclocking Error}
\subsubsection{Generation of Overclocking Error}\label{subsub:Generation of Overclocking Error}
For a given $T_S$, the maximum length of error-free carry propagation is described by~(\ref{Max.length of carry chain}), where $f_S$ denotes the sampling frequency.
%
%\vspace{-1ex}
\begin{eqnarray}\label{Max.length of carry chain}
%\small
%\footnotesize
%\scriptsize
    b := \left\lceil \frac{T_S}{\mu} \right\rceil= \left\lceil \frac{1}{\mu\cdot f_S}\right\rceil
\end{eqnarray}
%\normalsize

@@

$S_i$ and $\hat{S_i}$ can be determined using the equations from the previous statements in Section~\ref{subsub:Generation of Overclocking Error}. In the error-free case, the carry will propagate from bit $S_t$ to bit $S_{t+m-1}$, and we will obtain $S_{t+b}=S_{t+b+1}=\cdots=S_{t+m-2}=0$ for carry propagation, and $S_{t+m-1}=1$ for carry annihilation. However, when a timing violation occurs, the carry will not propagate through all these bits. Substituting these values into (\ref{Output Error}) yields (\ref{etm}). Interestingly, the value of overclocking error has no dependence on the length of carry chain $m$.
%
%Instead, $\hat{S}_{t+b}=\hat{S}_{t+b+1}=\cdots=\hat{S}_{t+m-2}=1$ and $\hat{S}_{t+m-1}=0$. It is worth noting that these expressions are obtained based on the assumption that all the internal bits are initialized to zero.
%
\begin{eqnarray}\label{etm}
%\small
%\footnotesize
  \begin{split}
    e_{tm} &= \frac{\left|2^{t+m-1}-2^{t+m-2}-\dots-2^{t+b}\right|}{2^n}\\
    &= 2^{t+b-n}
  \end{split}
\end{eqnarray}
%\normalsize

@@

\subsection{Area Overhead}
We then compare different adder structures in terms of the maximum word-length with respect to different operating frequencies and the corresponding area consumption. Fig.~\ref{CSA3stage Timing} illustrates the maximum word-lengths for RCA and CSA with 2 stages and 3 stages under a range of operating frequencies. We only investigate 3 stages as the maximum stages number predicted in (\ref{CSA_StageNumber}) is 3. However, similar analysis can be performed on CSA with any stage numbers and the overall conclusion will still hold. It can be seen from Fig.~\ref{CSA3stage Timing} that in comparison to RCA, CSA achieves greater word-length when frequency is initially increased. This corresponds to smaller truncation errors. RCA only outperforms than CSA when very high frequency is applied. In addition, the word-length of 3-stage CSA is always greater than 2-stage CSA across the entire frequency domain, as expected.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Timing.eps}
%  \caption{Maximum word-length for RCA and CSA across a variety of frequencies.}
%  \label{CSA3stage Timing}
%\end{figure}

@@

However, the accuracy benefits brought by CSA comes with the cost of large area overhead. Fig.~\ref{CSA3stage Area} depicts the number of Look-Up Tables (LUTs) in the FPGA used for all three structures. It can be seen that in order to meet a given frequency, the 3-stage CSA consumes $2.4\times\sim3.7\times$ area than RCA, while the number of the 2-stage CSA is $1.7\times\sim3.1\times$. This finding poses a question of whether CSA with higher stage numbers still offer better accuracy than RCA under a limited area budget.
%\begin{figure}[htbp]
%  \centering
%  \includegraphics[width=3.3in]{./Figures/CSA3stage_Area.eps}
%  \caption{Area overhead for RCA and CSA.}
%  \label{CSA3stage Area}
%\end{figure}

\subsection{Exploring Trade-offs Between Accuracy, Performance and Area}
To answer this question, the silicon area is incorporated into the trade-offs together with accuracy and performance. If the available hardware resources are limited, the full word-length of both CSA and RCA might not be implemented. The precision loss might results in large errors even at low frequencies. This is demonstrated by experiments where area constraints are applied besides the timing requirement. In the conventional design scenario, both of the constraints are met by reducing the word-length of the input signal. The errors at the outputs are recorded with the reference to the original word-length, which is 16 bits in the following experiments. In addition, we propose another circuit designing method where RCA is implemented with the maximum possible word-length under the given area budget, while the timing constraints are met by overclocking.

Our results could be of interest to a circuit designer in two aspects. On one hand, it is desirable to create a circuit that can run at a given frequency with the minimum achievable output errors and resource usage. On the other hand, the algorithm designer will wish the circuit to operate as fast as possible with the minimum area whilst a certain error budget can be tolerated. We select the optimum design methodology for both situations.

In the first case, for a certain pair of $\{Area, Frequency\}$ constraint, the optimum design method is selected based on the following criteria:
%For instance, Fig.xxx to Fig.xxx illustrate error expectations of all the described design scenarios when setting the LUTs number to 45, 35, 25 and 15 respectively. The optimal design metric with the minimum error expectation is labelled for each area constraint. For a given frequency and area constraint, the optimum design is chosen based on the following metrics:
\begin{itemize}
  \item Design with the best accuracy at outputs is the optimum design;
  \item If multiple designs achieve the same accuracy, then the design with minimum area is the optimum design;
  \item If the accuracy and area are identical for multiple designs, they are all treated as the optimum design.
\end{itemize}

The accuracy here is assessed in terms of the mean error at outputs. The optimum design method with respect to a variety of operating frequencies and area consumptions is demonstrated in Fig.~\ref{Tradeoff}, from which several observations can be made. If the available area is large enough, CSA serves as the optimum design option because it inherently operates faster than RCA, and it can be implemented with full precision with unlimited hardware resources. The 2-stage CSA is better than the 3-stage CSA when frequency is initially increased, as it consumes less area although both of them achieve the same error expectation. For a tighter area budget, only part of the CSA can be implemented, whilst RCA still keeps full precision. In this case, area instead of timing becomes the dominate factor. This leads to precision loss for CSA from the start. Therefore we see RCA with overclocking is the optimum design method across almost the whole frequency domain. For an even stringent area constraint, the word-length of RCA is also limited. This results in truncation error initially for all design scenarios. Meanwhile, CSA with high stage numbers might not even be implemented at all.
\begin{figure}[htbp]
  \centering
  \includegraphics[width=3.5in]{./Figures/Tradeoff.eps}
  \caption{Demonstration of the optimum design methodology which achieves the minimum error at outputs with respect to a variety of frequency and area constraints.}
  \label{Tradeoff}
\end{figure}

Similarly in the second case where a pair of $\{Area, Accuracy\}$ constraint is applied, a certain amount of error can be tolerated when designing circuits. The error specifications are evaluated in terms of mean relative error (MRE), as presented in~(\ref{MRE}), where $E_{error}$ and $E_{out}$ refer to the mean value of error and the mean value of outputs, respectively.
\begin{eqnarray}\label{MRE}
  MRE=\left|\frac{E_{error}}{E_{out}}\right|\times 100\%
\end{eqnarray}

In our experiments, MRE is set ranging from $0.001\%$ to $50\%$. For a certain MRE, design with the maximum operating frequencies is labeled as the optimum design. Moreover, the smallest design is the optimum one if multiple structures operate at the same frequency, with a certain area. Based on these criteria, the decision graph is depicted in Fig.~(\ref{Tradeoff_Error}).
%We also notice that for small frequencies, RCA with either overclocking or truncation of word-length of the input signal is the best design choice for any area constraint, since
\begin{figure}[htbp]
    \centering
    \includegraphics[width=3.5in]{./Figures/Tradeoff_Error.eps}
    \caption{Demonstration of the optimum design methodology which runs at the fastest frequency with respect to a variety of accuracy and area constraints.}
    \label{Tradeoff_Error}
\end{figure}

For a tight accuracy requirement, i.e. $MRE<0.005\%$, CSA serves as the optimum design choice with respect to large accessible area, as it intrinsically operates faster than RCA. When the area budget shrinks, RCA exceeds because the precision is CSA is limited. Similarly the precision of RCA is reduced when critical area requirement is applied.

However, if the tolerable error increases, i.e. $MRE>1\%$, the overclocked RCA achieves the fastest operating frequencies with respect to most area constraints.
